A Two-Infall Star Formation History with Outflows and Radial Migration

### 2025-02-18 src/tex/ms.tex
Updates from Overleaf

\begin{table*}
    \centering
    \caption{Sample selection parameters and median uncertainties from APOGEE DR17 (see Section \ref{sec:observational-sample}).}
    \label{tab:sample}
    \begin{tabular}{lll}
        \hline\hline
        \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Range or Value} & \multicolumn{1}{c}{Notes} \\
        \hline
        $\log g$            & $1.0 < \log g < 3.8$          & Select giants only \\
        $T_{\rm eff}$       & $3500 < T_{\rm eff} < 5500$ K & Reliable temperature range \\
        $S/N$               & $S/N > 80$                    & Required for accurate stellar parameters \\
        ASPCAPFLAG Bits     & $\notin$ 23                   & Remove stars flagged as bad \\
        EXTRATARG Bits      & $\notin$ 0, 1, 2, 3, or 4     & Select main red star sample only \\
        Age                 & $\sigma_{\rm Age} < 40\%$     & Age uncertainty from \citetalias{leung_variational_2023} \\
        $R_{\rm gal}$     & $3 < R_{\rm gal} < 15\,{\rm kpc}$    & Eliminate bulge \& extreme outer-disk stars \\
        $|z|$               & $|z| < 2\,{\rm kpc}$                 & Eliminate halo stars \\
        \hline
    \end{tabular}
\end{table*}

@@

\begin{table*}
    \centering
    \caption{Median and dispersion in APOGEE parameter uncertainties.}
    \label{tab:uncertainties}
    \begin{tabular}{lcc}
        \hline\hline
        \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Median Uncertainty} & \multicolumn{1}{c}{Uncertainty Dispersion ($95\%-5\%$)} \\
        \hline
        [Fe/H]          & $0.0089$   & $0.0060$ \\
        ${\rm [O/Fe]}$  & $0.019$    & $0.031$ \\
        log(Age/Gyr)    & $0.10$     & $0.16$ \\
        \hline
    \end{tabular}
\end{table*}

@@

We obtain a sample of \input{output/sample_size.txt} red giant branch and red clump stars with high-quality spectra using the selection criteria listed in Table \ref{tab:sample}, which are adapted from \citet{hayden_chemical_2015}. We use the \citet{bailer-jones_estimating_2021} photo-geometric distance estimates from {\it Gaia} Early Data Release 3 \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2021} included in the APOGEE DR17 catalog. When calculating the galactocentric radius $R_{\rm gal}$ and midplane distance $z$ of each star, we assume the distance between the Sun and the Galactic center is $R_\odot=8.122\,{\rm kpc}$ \citep{gravity_collaboration_detection_2018} and the Solar midplane distance is $z_\odot=20.8\,{\rm pc}$ \citep{bennett_vertical_2019}. Table \ref{tab:apogee-regions} presents the number of APOGEE stars in bins of $R_{\rm gal}$ and $|z|$; typical distance uncertainties are much smaller than the bin width. \todo{Mention calibrated abundances.}

We supplement the APOGEE DR17 abundance data with two different age catalogs.

\todo{Taken from last paper.} We use estimated ages from \citet[][hereafter \citetalias{leung_variational_2023}]{leung_variational_2023}, who use a variational encoder-decoder network which is trained on asteroseismic data to retrieve age estimates for APOGEE giants without contamination from age-abundance correlations. Importantly, the \citetalias{leung_variational_2023} ages do not plateau beyond $\sim10\,{\rm Gyr}$ as they do in astroNN \citep{mackereth_dynamical_2019}. We use an age uncertainty cut of 40\% per the recommendations of \citetalias{leung_variational_2023}, which produces a total sample of \input{output/age_sample_size.txt} APOGEE stars with age estimates. We note that we use the full sample of \input{output/sample_size.txt} APOGEE stars unless we explicitly compare to age estimates. Table \ref{tab:uncertainties} presents the median and dispersion ($95^{\rm th} - 5^{\rm th}$ percentile difference) of the uncertainty in [Fe/H], [O/Fe], and log(age).

>>>>

\begin{table*}
    \centering
    \caption{Sample selection parameters and median uncertainties from APOGEE DR17 (see Section \ref{sec:observational-sample}).}
    \label{tab:sample}
    \begin{tabular}{lll}
        \hline\hline
        \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Range or Value} & \multicolumn{1}{c}{Notes} \\
        \hline
        $\log g$            & $1.0 < \log g < 3.8$          & Select giants only \\
        $T_{\rm eff}$       & $3500 < T_{\rm eff} < 5500$ K & Reliable temperature range \\
        $S/N$               & $S/N > 80$                    & Required for accurate stellar parameters \\
        ASPCAPFLAG Bits     & $\notin$ 23                   & Remove stars flagged as bad \\
        EXTRATARG Bits      & $\notin$ 0, 1, 2, 3, or 4     & Select main red star sample only \\
        Latent age          & $\sigma_\tau/\tau < 40\%$     & Age uncertainty from \citet{leung_variational_2023} \\
        $R_{\rm gal}$     & $3 < R_{\rm gal} < 15\,{\rm kpc}$    & Eliminate bulge \& extreme outer-disk stars \\
        $|z|$               & $|z| < 2\,{\rm kpc}$                 & Eliminate halo stars \\
        \hline
    \end{tabular}
\end{table*}

@@

\begin{table*}
    \centering
    \caption{Median and dispersion in APOGEE parameter uncertainties.}
    \label{tab:uncertainties}
    \begin{tabular}{lcc}
        \hline\hline
        \multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{Median Uncertainty} & \multicolumn{1}{c}{Uncertainty Dispersion ($95\%-5\%$)} \\
        \hline
        [Fe/H]          & $0.0089$   & $0.0060$ \\
        ${\rm [O/Fe]}$  & $0.019$    & $0.031$ \\
        $\tau_{\rm latent}/{\rm Gyr}$    & 1.0   & $0.16$ \\
        $\tau_{\rm [C/N]}/{\rm Gyr}$     & 1.0   & \todo{N/A} \\
        \hline
    \end{tabular}
\end{table*}

@@

We obtain a sample of \num{171635} red giant branch and red clump stars with high-quality spectra using the selection criteria listed in Table \ref{tab:sample}, which are adapted from \citet{hayden_chemical_2015}. Table \ref{tab:uncertainties} presents the median statistical uncertainty and uncertainty dispersion ($95^{\rm th} - 5^{\rm th}$ percentile difference) of the calibrated [Fe/H] and [O/Fe] abundances for our sample. We use the \citet{bailer-jones_estimating_2021} photo-geometric distance estimates from {\it Gaia} Early Data Release 3 \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2021} included in the APOGEE DR17 catalog. When calculating the galactocentric radius $R_{\rm gal}$ and midplane distance $z$ of each star, we assume the distance between the Sun and the Galactic center is $R_\odot=8.122\,{\rm kpc}$ \citep{gravity_collaboration_detection_2018} and the Solar midplane distance is $z_\odot=20.8\,{\rm pc}$ \citep{bennett_vertical_2019}. Table \ref{tab:apogee-regions} presents the number of APOGEE stars in bins of $R_{\rm gal}$ and $|z|$; typical distance uncertainties are much smaller than the bin width.

We supplement the APOGEE DR17 abundance data with two different age catalogs. The first is the ``latent'' age estimates from \citet{leung_variational_2023}, who train a variational encoder-decoder network on asteroseismic data for APOGEE red giants. This catalog has two main advantages for our purposes: their method is designed to reduce contamination from abundance information (in particular [$\alpha$/Fe]), and the recovered ages do not plateau at $\sim10\,{\rm Gyr}$ as they do in some other neural network-derived age catalogs \citep[e.g.,][]{mackereth_dynamical_2019}. Following the recommendations of \citet{leung_variational_2023}, we use an upper limit on the relative age uncertainty of 40\%. This produces a sample of \num{57607} stars with latent age estimates, of which \num{14871} are in the Solar neighborhood ($7\leq R_{\rm gal}<9\,{\rm kpc}$, $0\leq|z|<0.5\,{\rm kpc}$). The median uncertainty in $\log({\rm age}/{\rm Gyr})$ is 0.10 (see Table \ref{tab:uncertainties}).

Our second age catalog uses the [C/N]--age relation for red giant branch and red clump stars. 
\todo{Paragraph describing [C/N]-derived ages (get help from Jack).}
This method has the benefit of providing age estimates for luminous giants, which increases the sample size at farther distances from the Sun. However, systematic effects of the [C/N]--age relation mean that these ages are most trustworthy in the range $1\sim10$ Gyr.
